---
title: "Group Assignment 2"
author: "Wahib Charamand, Valentin de la Pena, Jinwen Wu, Abe Rubin"
subtitle: "Analytics for Strategy"
date: "`r Sys.Date()`"
output: html_document
---

```{r, message=FALSE, warning=FALSE}
# Packages
library(dplyr)
library(lmtest)
library(sandwich)

# Load Dataset
united <- read.csv("/Users/jinwens/Downloads/united-1.csv")
```

```{r, message=FALSE, warning=FALSE}
# Data Cleaning
united <- united %>%
  mutate(
    bookingcountry_us = ifelse(bookingcountry == "US", 1, 0)
  )
```

```{r}
# Create Data Subset for Before Memorial Day Weekend
united_pre_memorial_day <- united_memorial_day <- united[!(united$departuredate %in% c("26-May-07", "27-May-07", "28-May-07")),]

# Create Logit Model
noshows_logit <- glm(noshow ~
                      numberinparty +
                      daystodep + 
                      bookingcountry_us +
                      eticketed + 
                      as.factor(assignedseat) +
                      upline + 
                      downline,
                      data = united_pre_memorial_day,
                      family=binomial(link = "logit") )
coeftest(noshows_logit, vcovCL, type='HC1')

# Calculate Baseline Odds

  # Choose Coefficient from Logit
noshows_logit_coef <- coef(noshows_logit)

  # Calculate Odds from Coefficient
baselineodds <- exp(noshows_logit_coef[1])

  # Print Baseline Odds
cat("Baseline Odds:", baselineodds, "\n")

  # Calculate Probability from Odds
baselineprobability <- (baselineodds)/(1+baselineodds)

  # Print Baseline Probability
cat("Baseline Probability:", baselineprobability, "\n")

# Calculate Odds of No-Show for "Typical" Business Traveler (Described Below)
  
  # Create Data for Hypothetical "Typical" Business Traveler
typical_business_traveler <- list(
  numberinparty = 1,
  daystodep = 7,
  bookingcountry_us = 1,
  eticketed = 1,
  assignedseat = "Y",
  upline = 0,
  downline = 0
)

  # Predict Result for Business Traveler
predicted_business_traveler <- predict(noshows_logit, 
                                           newdata = typical_business_traveler, 
                                           type = "response")

  # Print Probability for Business Traveler
cat("Baseline Probability for Business Traveler:", 
    predicted_business_traveler, "\n")

  # Create Data for Hypothetical "Typical" Business Traveler with Upline Connection
typical_business_traveler_with_upline <- list(
  numberinparty = 1,
  daystodep = 7,
  bookingcountry_us = 1,
  eticketed = 1,
  assignedseat = "Y",
  upline = 1,
  downline = 0
)

  # Predict Result for Business Traveler with Upline
predicted_business_traveler_with_upline <- predict(noshows_logit, 
                                           newdata = typical_business_traveler_with_upline, 
                                           type = "response")

  # Print Probability for Business Traveler
cat("Probability for Business Traveler with Upline:", 
    predicted_business_traveler_with_upline, "\n")
```

Question 1.

(a) The intercept in the above regression was -1.17790157. This implies an odds of 0.3079242 and probability of 0.2354297 of a no-show for a person when the predictor variables are at either 0 or their reference levels. Specifically, this is a person with 0 people in their party who booked or modified with 0 days to departure, made their booking not in the US as a non-eticket, without an assigned seat, and with no connections. Of course, this person does not make logical sense, as it is impossible to have 0 people in a party.

Question 2.

(a) The effect of adding an upline connection is statistically significant, however, the effect of adding a downline connection is not statistically significant.

(b) The regression coefficient on the upline connection was 0.90617180. This implies that having an upline connection multiplies an individual's odds of no-showing by 0.90617180. This means that the odds of no-showing decrease.

(c) We do not know the implied effect of an upline connection on the party's probability of a no-show without having a baseline probability. For a baseline probability, consider an individual who is traveling alone, books their flight 7 days before departure from within the U.S. as an e-ticket with an assigned seat and no connections. This baseline represents what might be a typical business traveler, so, they make sense to use as a baseline for understanding the impact of a connection. The probability of a no-show for this traveler is 0.05297631. However, when this traveler has an upline connection, the probability of a no-show becomes 0.1216062. So, adding an upline connection for what one may consider the "typical" business traveler increases the probability of a no show by approximately 7%.

```{r}
# Create Data Subset for Memorial Day Weekend
united_memorial_day <- united[(united$departuredate %in% c("26-May-07", "27-May-07", "28-May-07")),]

# Calculate No Shows Probability
probability_noshows_memorial_day <- predict(noshows_logit, united_memorial_day, type = "response")
united_memorial_day$predicted_noshows_prob <- probability_noshows_memorial_day

# Calculate Predicted No Shows for Party
united_memorial_day$predicted_noshows_memorial_day <- united_memorial_day$numberinparty*united_memorial_day$predicted_noshows_prob

# Calculate Actual No Shows for a Party (Total)
united_memorial_day$actual_noshows <- united_memorial_day$numberinparty*united_memorial_day$noshow

# Examine Predicted No-Shows vs. Realized
cat("Actual No-Shows", 
    sum(united_memorial_day$actual_noshows), "\n")
cat("Predicted No-Shows", 
    sum(united_memorial_day$predicted_noshows_memorial_day), "\n")

```

Question 3:

(a) Our predicted number of no-shows was 64.58884, compared to 45 actual no-shows over Memorial Day Weekend. Therefore, we slightly overestimated the number of no-shows. This would not be good for United, as, overestimating the number of no-shows could result in more passengers showing up than are seats available on flights.

(b) 
