---
title: "Individual assignment 2, Suggested answers"
author: "Shengmaof Cao"
output:
  html_document: default
---

```{r}
#install.packages("modelsummary")
library(modelsummary) 
```

Q1

The key is to recognize default is a binary outcome variable.
We can use the binary outcome sample size calculation formula from class.

```{r}
#type1 = 0.05
a = 1.96
#type2 = 0.8
b = 0.84
p1 = 0.27
p2 = 0.09

pbar = (p1 + p2) / 2
numerator = 2 * ((a + b)**2)*pbar*(1-pbar)
denominator = (p1 - p2)**2
N = numerator / denominator
print(N)

```

We need a sample size of 72 on each arm of the randomized trial 
to achieve the desired certainty on this measurement, totaling 144 observations.
The sample of 845 borrowers is large enough for our purposes.

------------------------------------------------------------------------

Q2)

We need information on intra-cluster corrleation and inter-cluster correlation to 
determine whether we have enough sample size. In the extreme case where everyone 
within a cluster is the same and all of the clusers are different, we would have 
effectively 169 observations which is above the 144 required for the RCT. 

------------------------------------------------------------------------

3)

```{r}
microfinance <- read.csv('./microfinance.csv')
str(microfinance)
```

```{r}
microfinance_pre <- subset(microfinance, select= 
    c(GracePeriod, loanamount, Age, Married, Years_Education, 
      HH_Size, Shock_Any, Has_Business, Homeowner))

#microfinance_pre$clusters <- microfinance$group_id
datasummary_balance(~GracePeriod, data = microfinance_pre, 
                    dinm_statistic= "p.value", fmt=3)
```


We observe that the randomization is in large part successful. "Married" is the 
only variable with an individual p-value below 0.05, but it is still higher than 
a BH-corrected threshold of 0.00625. We are comfortable with going ahead with 
the experiment. 
